### 2018-06-06 correspondence/180606-lecoanet/brief.tex
final version of code

The domain of our simulation is $x \in [0, H], z \in [0, 4H]$. We use a Fourier
basis in the $x$ and a Chebyshev in the $z$. We try with just $16$ modes in the
$x$ and $64$ modes in the $z$. Since we choose parameters of the problem such
that $\frac{2\pi}{k_x} = H, \frac{2\pi}{k_z} = -\frac{4H}{\pi}$\footnote{As
seen later, $k_z$ never explicitly enters into the problem description; instead,
we specify $\omega(k_x, k_z)$ using the desired $k_x, k_z$ values.}, this should
be plenty of resolution to resolve the IGW (I chose $k_z$ as an irrational value
to ensure any undesired reflections are far from any normal modes of the box).
The timescale is set such that $N = 1$.

@@

To simulate a purely outgoing wave, we adopt a damping zone near the top
boundary to suppress reflection. Specifically, we use
\begin{subequations}\label{eq:lin_feq_sponge}
    \begin{align}
        \vec{\nabla} \cdot \vec{u}_1 &= 0,\\
        \pd{\rho_1}{t} + f(z)\rho_1 - u_{1z}\frac{\rho_0}{H} &= 0,\\
        \pd{\vec{u}_1}{t} + f(z)\rho_1 + \frac{\vec{\nabla}P_1}{\rho_0}
            + \frac{\rho_1}{\rho_0}\vec{g} &= 0,
    \end{align}
\end{subequations}
where
\begin{equation}
    f(z) = f_0 \frac{\max\p*{z - z_0, 0}^2}{\p*{z_{\max} - z_0}^2}.
\end{equation}
We use $f_0 = 2, z_0 = 0.7 z_{\max}$, so the damping turns on quadratically
starting at $z_0 = 0.7z_{\max}$.

@@

To excite the wave, we adopt bottom boundary condition (BC)
\begin{equation}
    \at{\pd{u_{1z}}{z}}_{z = 0} = -Ak_z\cos\p*{k_xx - \omega t + \frac{1}{2H}}.
\end{equation}
The motivation for this comes from differentiating $\pd{u_{1z}}{z}$ in
\autoref{eq:lin_sol}. The reason our BC isn't just $u_{1z}(z = 0)$ is because the
initial condition $u_x(t = 0) = u_{1z}(t = 0) = 0$ is not divergence free.

@@

A snapshot of the simulation produced by \texttt{strat.py} is shown below in
\autoref{fig:lin_plot}. We see reasonable agreement between the analytical
solution and the simulation data as the wave propagates. More plots can be found
in the \texttt{plots/} folder.
\begin{figure}[t]
    \centering
    \begin{subfigure}{0.7\textwidth}
        \centering
        \includegraphics[width=\textwidth]{plots/t_20.png}
    \end{subfigure}

    \begin{subfigure}{0.7\textwidth}
        \centering
        \includegraphics[width=\textwidth]{plots/t_70.png}
    \end{subfigure}
    \caption{For line plots, red line indicates beginning of damping zone, blue
    the analytical solution and orange the simulation data. Variables are
    $u_{1z} = uz, u_{1x} = ux$, $F_z$ the energy flux, $E$ the energy, $P_1 = P,
    \rho_1 = rho$. Line plots of dynamical variables are slices through $x = 0$.
    }\label{fig:lin_plot}
\end{figure}

@@

To move to the nonlinear problem, we first introduce the nonlinear terms as in
\autoref{eq:nonlin_feq} but use small $A$ such that $u_{1z} \sim
Ae^{z_{\max}/2H} \ll v_{ph}$. All other problem setup is kept the same. The code
is in \texttt{strat\_nonlin\_lowA.py}. We expect agreement with the analytical
solution to the linear problem.

@@

We now increase the amplitude, and a numerical instability seems to develop. We
explore this with \texttt{strat\_nonlin.py}. An instability seems to grow near
$z = 0$.

@@

It seems that the observed instability has to do with the interface driving. A
few possible solutions have been considered:
\begin{itemize}
    \item I've considered other types of boundary conditions, such as
        $\ptd{u_{1z}}{z}$ or $\pd{P}{z}$, but all of these seem to produce
        similar behavior, instability near $z = 0$.

>>>>

The domain of our simulation is $x \in [0, H], z \in [0, 4H]$. We use a Fourier
basis in the $x$ and a Chebyshev in the $z$. We try with just $16$ modes in the
$x$ and $64$ modes in the $z$. Since we choose parameters of the problem such
that $\frac{2\pi}{k_x} = H, \frac{2\pi}{k_z} = -\frac{4H}{\pi}$\footnote{As
seen later, $k_z$ never explicitly enters into the problem description; instead,
we specify $\omega(k_x, k_z)$ using the desired $k_x, k_z$ values.}, this should
be plenty of resolution to resolve the IGW\footnote{I chose $k_z$ as an
irrational value to ensure any undesired reflections are far from any normal
modes of the box.}. The timescale is set by $N = 1$.

@@

To simulate a purely outgoing wave, we adopt a damping zone near the top
boundary to suppress reflection. Specifically, we use
\begin{subequations}\label{eq:lin_feq_sponge}
    \begin{align}
        \vec{\nabla} \cdot \vec{u}_1 &= 0,\\
        \pd{\rho_1}{t} + f(z)\rho_1 - u_{1z}\frac{\rho_0}{H} &= 0,\\
        \pd{\vec{u}_1}{t} + f(z)\rho_1 + \frac{\vec{\nabla}P_1}{\rho_0}
            + \frac{\rho_1}{\rho_0}\vec{g} &= 0,
    \end{align}
\end{subequations}
where
\begin{equation}
    f(z) = f_0 \frac{\max\p*{z - z_0, 0}^2}{\p*{z_{\max} - z_0}^2}.
\end{equation}
We use $f_0 = 1, z_0 = 0.7 z_{\max}$, so the damping turns on quadratically
starting at $z_0 = 0.7z_{\max}$.

@@

To excite the wave, we adopt bottom boundary condition (BC)
\begin{equation}
    \at{\pd{u_{1z}}{z}}_{z = 0} = -Ak_z\cos\p*{k_xx - \omega t + \frac{1}{2H}}.
\end{equation}
We use $A = 0.05$. The reason our BC isn't just $u_{1z}(z = 0)$ is because the
initial condition $u_x(t = 0) = u_{1z}(t = 0) = 0$ is not divergence free and
produces high-$k$ modes in $u_x$. We just differentiate $\pd{u_{1z}}{z}$ in
\autoref{eq:lin_sol} and take real part.

@@

A snapshot of the simulation produced by \texttt{strat.py} is shown below in
\autoref{fig:lin_plot}. We see reasonable agreement between the analytical
solution and the simulation data as the wave propagates. More plots can be found
in the \texttt{plots/} folder.
\begin{figure}[t]
    \centering
    \begin{subfigure}{0.75\textwidth}
        \centering
        \includegraphics[width=\textwidth]{plots/t_20.png}
    \end{subfigure}
    \begin{subfigure}{0.75\textwidth}
        \centering
        \includegraphics[width=\textwidth]{plots/t_70.png}
    \end{subfigure}
    \caption{For line plots, red line indicates beginning of damping zone, blue
    the analytical solution and orange the simulation data. Variables are
    $u_{1z} = uz, u_{1x} = ux$, $F_z$ the energy flux, $E$ the energy, $P_1 = P,
    \rho_1 = rho$. Line plots of dynamical variables are slices through $x = 0$.
    }\label{fig:lin_plot}
\end{figure}

@@

To move to the nonlinear problem, we first introduce the nonlinear terms as in
\autoref{eq:nonlin_feq} but use small $A$ such that $u_{1z} \sim
Ae^{z_{\max}/2H} \ll v_{ph}$; we use $A = \scinot{5}{-5}$. All other problem
setup is kept the same. The code is in \texttt{strat\_nonlin\_lowA.py}. We
expect agreement with the analytical solution to the linear problem, and this is
shown well in \autoref{fig:nonlin_plot_lowA}. Further plots are available in
\texttt{plots\_nonlin\_lowA/}.
\begin{figure}[t]
    \centering
    \begin{subfigure}{0.75\textwidth}
        \centering
        \includegraphics[width=\textwidth]{plots_nonlin_lowA/t_20.png}
    \end{subfigure}
    \begin{subfigure}{0.75\textwidth}
        \centering
        \includegraphics[width=\textwidth]{plots_nonlin_lowA/t_70.png}
    \end{subfigure}
    \caption{Same plots as \autoref{fig:lin_plot} but with nonlinear terms added
    back to the fluid equations and very small $A = \scinot{5}{-5}$.
    }\label{fig:nonlin_plot_lowA}
\end{figure}

@@

We now increase the amplitude back to $A = 0.05$, and a numerical instability
seems to develop. We explore this with \texttt{strat\_nonlin.py}. An instability
seems to grow near $z = 0$, and two plots of its evolution can be seen in
\autoref{fig:nonlin_plot}. Further plots can be found in
\texttt{plots\_nonlin/}.
\begin{figure}[t]
    \centering
    \begin{subfigure}{0.75\textwidth}
        \centering
        \includegraphics[width=\textwidth]{plots_nonlin/t_70.png}
    \end{subfigure}
    \begin{subfigure}{0.75\textwidth}
        \centering
        \includegraphics[width=\textwidth]{plots_nonlin/t_77.png}
    \end{subfigure}
    \caption{Nonlinear equations with $A = 0.05$ as it diverges near $z = 0$.
    }\label{fig:nonlin_plot}
\end{figure}

@@

It seems that the observed instability has to do with the interface driving. A
few possible solutions have been considered:
\begin{itemize}
    \item Doubling the temporal resolution doesn't seem to change when the
        instability sets in. Doubling the spatial resolution makes it set in
        earlier, and halving makes it set in later.

    \item I've considered other types of boundary conditions, such as
        $\ptd{u_{1z}}{z}$ or $\pd{P}{z}$, but all of these seem to produce
        similar behavior, instability near $z = 0$.
